MGRFE: multilayer recursive feature elimination based on an embedded genetic algorithm for cancer classification

### 2019-06-23 MGRFE-TCBB-2019-06-03/MGRFE-tcbb.tex
Keep Calm and Carry On.

	Filter algorithms generally evaluate features according to the inherent characteristic of a dataset, then rank all the features and preserve only an optimal subset of the original features. Up to now, lots of filter algorithms have been designed, such as the methods based on the \textit{t}-test \cite{W23}, $\chi^2$ test \cite{W21}, mutual information \cite{W22}, maximal information coefficient (MIC) \cite{W25}, and signal-to-noise ratio \cite{W24}. The \textit{t}-test is the frequently used and efficient statistical approach to detecting differentially expressed genes in microarray analysis \cite{RN413, RN414, RN421, RN417, RN418, RN419, RN416}. MIC is an information theory-based measurement for capturing a wide range of associations, which has shown excellent performance on detecting novel associations in large datasets \cite{W8}. The recent study of McTwo \cite{W16}, which is based on MIC for selection of a gene subset in a microarray data, has outperformed most of existing algorithms. In addition, MIC may offer more convenience in dealing with multiclass datasets. Hence, the \textit{t}-test and MIC are practicable and qualified approaches for selecting statistically significant discriminative genes and thus are mostly used in the feature preprocessing stage to generate a candidate gene set in the analysis of microarray. Because there is no classification algorithm involved in the filter algorithm, its computational speed is high and suitable for large datasets. On the other hand, the filter techniques for gene feature selection also have some limitations. First, filter methods are likely to add redundant features into the chosen subsets, which will lead to inaccessibility of optimal results. Second, the mutual information between features is ignored for the implicit orthogonality assumption of features \cite{P1}. Third, the features top-ranked by a filter algorithm are not always the best features for classification \cite{W16}.

	Wrapper algorithms usually employ classification models and contain heuristic rules to select feature subsets guided by the classification performance on the feature subsets being used, which are usually superior to filter algorithms but more time-consuming. A variety of wrapper algorithms have emerged involving simulated annealing, randomized hill climbing \cite{W26}, regularized random forest (RRF) \cite{W11}, particle swarm optimization (PSO) \cite{W12,W10}, and genetic algorithm (GA) \cite{W13}. With the rapid development of heuristic rules and evolutionary strategies commonly present in wrapper techniques, various swarm intelligence algorithms have been applied to the optimization of feature selection. Kar \emph{et~al.} have proposed a particle swarm optimization method based on adaptive K-nearest neighborhood (KNN) to identify a minimum meaningful gene subset \cite{W15}. Moosa \emph{et~al.} have presented a modified artificial bee colony algorithm (ABC) to select a minimum number of genes with high predictive accuracy for cancer classification \cite{P32}. Oreski \emph{et~al.} have designed a hybrid GA with neural networks to identify an optimal feature subset with high classification accuracy and scalability for credit risk assessment \cite{W17}. Jung and Zscheischler have described a guided hybrid GA to minimize the number of cost function evaluations \cite{W18}. Nevertheless, all these feature selection methods based on swarm intelligence algorithms use the binary encoding method and lack an explicit reduction in the feature number. The feature number only changes in the randomized evolution operation like mutation and crossover. Thus, these methods lack the precise control over the gene features in the individuals and can not explicitly remove genes to decrease the feature number. Meanwhile, it has been verified that only a minimal number of informative genes is enough for effective diagnosis of different phenotypes in microarray gene datasets \cite{P1,W16,P32,P33}. The feature selection using binary encoding has three main shortcomings in finding an optimal gene combination: (1) The fixed chromosome length for the encoding length must be equal to the gene range to represent all the genes. This arrangement can result in impossibility of the explicit reduction in the gene number and unnecessary space occupation when there are only several 1s among lots of 0s. (2) There are different numbers of actual existing genes in different individuals. Because there are different numbers of 1s, the actual number of genes varies among individuals and cannot be controlled precisely. (3) The convergence speed is usually low and the time cost is high to generate the minimal informative gene combination. 
	The sizes of the optimal gene combinations in most of datasets are usually below 10. The evolution-based feature selection algorithms using binary encoding lack of an explicit feature reduction mechanism, which results in low probability and high time cost to generate the optimal minimal gene combination among the several thousand to tens of thousands of genes in each dataset.
	Recursive feature elimination (RFE) is a popular strategy that yields an explicit recursive feature reduction by recursively removing features with the least weights \cite{W7,P1,P2,P27}.

@@

	\begin{figure*}[!htbp]
		\centering
		\includegraphics[width=0.57\textwidth]{pictures/MGRFE-whole.pdf}
		\caption{A flowchart of the whole MGRFE procedure in this study.}
		\label{Fig:MGRFE_whole}
	\end{figure*}

@@

	This study involves 19 benchmark microarrays including binary, multiclass, balanced, and imbalanced datasets, which are subdivided into two large datasets.
	Dataset One consists of the 17 binary classification datasets used in ref. \cite{W16}, which includes diffuse large B-cell lymphoma (DLBCL) \cite{Y22}, Prostate (Pros) \cite{Y23}, acute lymphoblastic leukemia (ALL; subdivided into four subtypes based on different phenotypes) \cite{Y24}, central nervous system embryonal tumor (CNS) \cite{Y25}, Lymphoma (Lym) \cite{Y26}, Adenoma (Adeno) \cite{Y27}, Colon \cite{Y20}, Leukaemia (Leuk) \cite{Y21}, Myeloma (Mye)\cite{Y28}, Gastric (Gas) \cite{Y29}, and Gastric1/Gastric2 (Gas1/Gas2) cancer \cite{Y30} as well as type 1 diabetes (T1D) \cite{Y31}, and Stroke \cite{Y32}. Among them, DLBCL, Colon, Leukaemia, Myeloma, ALL1-4, and CNS datasets are imbalanced.
	Dataset Two is composed of the three typical benchmark microarray datasets used in ref. \cite{W15}, including two multiclass datasets of small-round blue-cell tumor (SRBCT) \cite{P13} and mixed lineage leukemia (MLL) \cite{W3} and one binary dataset of acute lymphoblastic leukemia and acute myeloid (ALL\_AML). Many previous experiments are conducted on these three datasets \cite{P8,P9,P10,P11,P12,P13,P14,P15,P16,P17,P18,P19,P20,P21,P22,P23}. The performance comparison between these methods and MGRFE will be provided in the Results section. It should be mentioned that the widely used benchmark Leukaemia was tested in both refs. \cite{W16} and \cite{W15} named as Leuk and ALL\_AML respectively, but they are same actually.
	The information on the 19 datasets in the two large datasets is given in Tables 1 and 2 in Supplementary Material. All these datasets can be downloaded directly from \url{https://github.com/Pengeace/MGRFE-GaRFE}.

@@

	On the 17 binary classification datasets, we performed six widely used measurements to compare the performance: Accuracy ($Acc$), Sensitivity ($Sn$), Specificity ($Sp$), Average accuracy ($Avc$), Matthews Correlation Coefficient($MCC$), and $AUC$. $AUC$ is the area under the receiver operating characteristic (ROC) curve, and the formulas of the other five measurements are presented in Equation (\ref{eq:1}). In Equation (\ref{eq:1}), $P$ and $N$ represent the numbers of positive and negative samples; $TP$ and $TN$ mean the number of correctly predicted positive and negative samples; and $FP$ and $FN$ denote the wrongly predicted positive and negative samples, respectively. For the two multiclass datasets, for consistency and convenience, only \(Acc\) is used.

@@

	As shown in Fig. \ref {Fig:MGRFE_flowchart}, the proposed MGRFE is divided into three stages, which are (1) Search space reduction; (2) Precise wrapper search and (3) Multiple k-fold cross-validation (CV). MGRFE is a multilayer iterative feature selection method with GA-RFE acting as the feature selection unit in every layer. GA-RFE is a recursive feature elimination process including embedded GA. The Pseudocodes 1, 2 and 3 in Supplementary Material provide the pseudocodes of the processes MGRFE, GA-RFE and embedded GA.

	\begin{figure*}[t]
		\centering
		\includegraphics[width=0.87\textwidth]{pictures/MGRFE-flowchart.pdf}
		\caption{The flowchart of MGRFE, which is divided into 3 stages: search space reduction, precise wrapper search, and multiple k-fold CV. Stage 2 is the core of MGRFE, which includes two key processes: GA-RFE and embedded modified GA.}
		\label{Fig:MGRFE_flowchart}
	\end{figure*}

@@

	At Stage 1, two filter methods, the \textit{t}-test and MIC, are used to decrease the gene range and offer a candidate gene set for later precise wrapper search stage. First, we perform the \textit{t}-test on all genes and subject them to ascending sorting according to their $p$ values, then the top-ranked statistically significant features with $p$ values less than 0.05 are preserved. Next, the upper limit of the features kept after the \textit{t}-test is set to 1000, that is, when there are more than 1000 features having $p$ values less than 0.05, only the top 1000 with lower $p$ values would be kept. If the preserved features after the \textit{t}-test screening are fewer than 500, they are all kept directly and definitively to form the candidate gene set without MIC screening; otherwise, the MIC-based selection will be followed. Second, we carry out MIC calculation on the preserved genes and resort them according to their MIC values, then the candidate gene set is generated from the top 500 genes with higher MIC values. For the two multiclass datasets, a candidate gene set is generated based only on the descending order of MIC values of all genes for which the multivariate \textit{t}-test cannot be performed directly. 
	In the Table 5 in Supplementary Material, the number of statistically significant genes with \emph{t}-test-based \emph{p}-values less than 0.05 on each binary-class dataset are listed.
	In the "S8" section of Supplementary Material, we also give a simple comparison of \textit{t}-test+MIC with other filter combinations.

@@

	\subsubsection{Stage 3: Multiple k-fold CV to select the final gene combination}

	Stage 3 is aimed at finding the optimal gene combination with the best classification performance and minimal variance among different CV processes. K-fold CV is used for calculating the fitness of a GA individual. Multiple k-fold CV based on different random seeds is performed to further validate and select the final optimal gene combination.

@@

	\begin{table*}[!htbp] \footnotesize
		\centering
		\caption{Results of MGRFE on 17 datasets in Dataset One}
		\label{Tab:RST all index}
		\resizebox{0.95\textwidth}{!}{
			\begin{tabular}{l|lllllllll}
				\toprule
				Datasets & Pos/Neg & Genes/Total & $Sn$ & $Sp$ & $Acc$ & $Avc$ & $MCC$ & $AUC$ & \textit{t}-test/MIC-based gene rankings\\
				\midrule
				DLBCL & 58/19 & \textbf{3}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [13/8, 39/24, 54/52]\\
				Pros & 52/50 & \textbf{4}/12625 & 0.980 & 0.982 & \textbf{0.981} & 0.981 & 0.963 & 0.98 & [1/1, 15/47, 74/49, 694/618]\\
				Colon & 40/22 & \textbf{6}/2000 & 1.0 & 0.960 & \textbf{0.985} & 0.980 & 0.969 & 0.97 & [15/6, 58/21, 176/297, 225/80, 240/555, 495/482]\\
				Leuk & 47/25 & \textbf{2}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [4/3, 7/5]\\
				Mye & 137/36 & \textbf{7}/12625 & 0.963 & 0.839 & \textbf{0.937} & 0.901 & 0.816 & 0.95 & [3/3, 15/103, 83/142, 143/13, 378/217, 404/644, 569/707]\\
				ALL1 & 95/33 & \textbf{1}/12625 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [1/1]\\
				ALL2 & 65/35 & \textbf{8}/12625 & 0.914 & 0.908 & \textbf{0.910} & 0.911 & 0.829 & 0.94 & [1/80, 52/395, 78/3040, 80/1297, 522/2448, 687/2038, 737/920, 760/1449]\\
				ALL3 & 24/101 & \textbf{8}/12625 & 0.830 & 0.950 & \textbf{0.927} & 0.890 & 0.785 & 0.93 & [4/500, 52/3437, 75/3010, 142/393, 488/443, 510/795, 715/1551, 770/1321]\\
				ALL4 & 26/67 & \textbf{6}/12625 & 1.0 & 0.986 & \textbf{0.990} & 0.993 & 0.978 & 0.99 & [1/2, 6/45, 39/356, 282/226, 535/497, 754/1377]\\
				CNS & 39/21 & \textbf{7}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 0.98 & [9/907, 53/542, 130/620, 131/519, 272/57, 273/454, 520/49]\\
				Lym & 22/23 & \textbf{3}/4026 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [4/7, 5/4, 669/135]\\
				Adeno & 18/18 & \textbf{1}/7457 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [468/27]\\
				Gas & 29/36 & \textbf{3}/22645 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [22/1, 77/32, 306/36]\\
				Gas1 & 72/72 & \textbf{3}/22283 & 0.986 & 0.973 & \textbf{0.980} & 0.980 & 0.961 & 0.99 & [132/74, 248/167, 717/500]\\
				Gas2 & 62/62 & \textbf{2}/22283 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [38/6, 89/62]\\
				T1D & 57/44 & \textbf{7}/54675 & 0.911 & 0.912 & \textbf{0.911} & 0.912 & 0.826 & 0.94 & [14/2229, 25/1579, 113/1287, 559/1282, 578/353, 680/426, 978/1728]\\
				Stroke & 20/20 & \textbf{4}/54675 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [1/3, 23/115, 129/543, 276/539]\\
				\bottomrule
		\end{tabular}}
	\end{table*}
	\begin{figure*}[htbp]
		\centering
		\includegraphics[width=0.72\textwidth]{pictures/CV1.pdf}
		\caption{The distribution of $Acc$ values in 10-time 10-fold CV for the selected gene combinations of 17 datasets in Dataset One.}
		\label{Fig:CV1}
	\end{figure*}

@@

	McTwo \cite{W16} thoroughly tested all the datasets in Dataset One and demonstrated satisfactory performance. Here, we present the performance comparison between McTwo and MGRFE. Table \ref{Tab:CMP McTwo RST} lists the overall maximal \(Acc\) and numbers of selected genes on total 17 datasets for MGRFE and McTwo. 
	On all 17 datasets, MGRFE can obtain equal or better prediction performance compared to McTwo. 
%	For a more intuitive comparison, Fig. \ref{Fig:CMP McTwo} offers the line chart of maximal \(Acc\) achieved by the two algorithms on 17 datasets, where MGRFE obviously outperforms McTwo with a higher \(Acc\) line. 	
	On five datasets ALL2, ALL3, ALL4, Stroke and CNS, MGRFE achieves distinctly better classification performance than McTwo with relatively more genes. For a fairer and more specific comparison, the \(Acc\) values of the two algorithms are listed when the gene number of MGRFE is equal to McTwo as shown in Table \ref{Tab:same gene num}. The results indicate that MGRFE still outperforms McTwo. Nonetheless, the \(Acc\) values associated with the usage of the gene numbers fall behind our optimal \(Acc\) values on these datasets. Thus, MGRFE selected somewhat more genes to achieve the optimal results.

%	\begin{figure*}[ht]
%		\centering
%		\includegraphics[width=0.72\textwidth]{pictures/CMP-McTwo.pdf}
%		\caption{The line plots of overall maximal accuracy for MGRFE and McTwo on 17 datasets in Dataset One.}
%		\label{Fig:CMP McTwo}
%	\end{figure*}

@@

%	\begin{figure*}[!htbp]
%		\begin{minipage}[t]{0.3\textwidth}
%			\centering
%			\includegraphics[height=1.55in]{pictures/SRBCT-RFE.pdf}
%		\end{minipage}
%		\begin{minipage}[t]{0.3\textwidth}
%			\centering
%			\includegraphics[height=1.55in]{pictures/ALLAML-RFE.pdf}
%		\end{minipage}
%		\begin{minipage}[t]{0.3\textwidth}
%			\centering
%			\includegraphics[height=1.55in]{pictures/MLL-RFE.pdf}
%		\end{minipage}
%		\caption{Three typical GA-RFE processes on the three benchmark datasets from Dataset Two.}
%		\label{Fig:3 RFE on benchmarks}
%	\end{figure*}

	\begin{figure}[htbp]
		\centering
		\includegraphics[width=0.45\textwidth]{pictures/CV2.pdf}
		\caption{The distribution of $Acc$ values in 10-time 10-fold CV for the selected gene combinations in Dataset Two.}
		\label{Fig:CV2}
	\end{figure}

@@

	\subsection{Independent validation of selected features}
	We performed totally 10-group validation experiments on independent datasets to verify the generalization ability of selected gene subsets by MGRFE. For each experiment, firstly, the selected gene probe features from the first dataset were transformed into the official gene symbols; secondly, the obtained gene symbols were transformed into corresponding gene probe Ids in the second dataset; thirdly, three kinds of classifier were used to perform 10 times \textit{k}-fold cross validation using the samples and selected gene probe features on the second dataset. Particularly, no feature mapping between Gas1 and Gas2 for they are generated simultaneously and have identical feature set.

>>>>

	Filter algorithms generally evaluate features according to the inherent characteristic of a dataset, then rank all the features and preserve only an optimal subset of the original features. Up to now, lots of filter algorithms have been designed, such as the methods based on the \textit{t}-test \cite{W23}, $\chi^2$ test \cite{W21}, mutual information \cite{W22}, maximal information coefficient (MIC) \cite{W25}, and signal-to-noise ratio \cite{W24}. The \textit{t}-test is the frequently used and efficient statistical approach to detecting differentially expressed genes in microarray analysis \cite{RN413, RN414, RN421, RN417, RN418, RN419, RN416}. MIC is an information theory-based measurement for capturing a wide range of associations, which has shown excellent performance on detecting novel associations in large datasets \cite{W8}. The recent study of McTwo \cite{W16}, which is based on MIC for selection of a gene subset in a microarray data, has outperformed most of existing algorithms. In addition, MIC may offer more convenience in dealing with multiclass datasets. Hence, the \textit{t}-test and MIC are practicable and qualified approaches for selecting statistically significant discriminative genes and thus are used in the feature preprocessing stage in this study to generate the candidate gene sets. Because there is no classification algorithm involved in the filter algorithm, its computational speed is high and suitable for large datasets. On the other hand, the filter techniques for gene feature selection also have some limitations. First, filter methods are likely to add redundant features into the chosen subsets, which will lead to inaccessibility of optimal results. Second, the mutual information between features is ignored for the implicit orthogonality assumption of features \cite{P1}. Third, the features top-ranked by a filter algorithm are not always the best features for classification \cite{W16}.

	Wrapper algorithms usually employ classification models and contain heuristic rules to select feature subsets guided by the classification performance on the feature subsets being used. These methods are usually superior to filter algorithms but more time-consuming. A variety of wrapper algorithms have emerged involving randomized hill climbing \cite{W26}, regularized random forest (RRF) \cite{W11}, particle swarm optimization (PSO) \cite{W12,W10}, and genetic algorithm (GA) \cite{W13}. With the rapid development of heuristic rules and evolutionary strategies commonly present in wrapper techniques, various swarm intelligence algorithms have been applied to the optimization of feature selection. Kar \emph{et~al.} have proposed a particle swarm optimization method based on adaptive K-nearest neighborhood (KNN) to identify a minimum meaningful gene subset \cite{W15}. Moosa \emph{et~al.} have presented a modified artificial bee colony algorithm (ABC) to select a minimum number of genes with high predictive accuracy for cancer classification \cite{P32}. Oreski \emph{et~al.} have designed a hybrid GA with neural networks to identify an optimal feature subset with high classification accuracy and scalability for credit risk assessment \cite{W17}. Jung and Zscheischler have described a guided hybrid GA to minimize the number of cost function evaluations \cite{W18}. Nevertheless, all these feature selection methods based on swarm intelligence algorithms use the binary encoding method and lack an explicit reduction in the feature number. The feature number only changes in the randomized evolution operators like mutation and crossover. Thus, these methods lack the precise control over the gene features in individuals and can not explicitly remove genes to decrease the feature number. Meanwhile, it has been verified that only a minimal number of informative genes is enough for effective diagnosis of different phenotypes in microarray gene datasets \cite{P1,W16,P32,P33}. The feature selection using binary encoding has three main shortcomings in finding an optimal gene combination: (1) The fixed chromosome length for the encoding length must be equal to the gene range to represent all the genes. This arrangement can result in impossibility of the explicit reduction in the gene number and unnecessary space occupation when there are only several 1s among lots of 0s. (2) There are different numbers of actual existing genes in different individuals. Because there are different numbers of 1s, the actual number of genes varies among individuals and cannot be controlled precisely. (3) The convergence speed is usually low and the time cost is high to generate the minimal informative gene combination. 
	The sizes of the optimal gene combinations in most of datasets are usually below 10. The evolution-based feature selection algorithms using binary encoding lack of an explicit feature reduction mechanism, which results in low probability and high time cost to generate the optimal minimal gene combination among the several thousand to tens of thousands of genes in each dataset.
	Recursive feature elimination (RFE) is a popular strategy that yields an explicit recursive feature reduction by recursively removing features with the least weights \cite{W7,P1,P2,P27}.

@@

	\begin{figure*}[!htbp]
		\centering
		\includegraphics[width=0.57\textwidth]{pictures/MGRFE-whole}
		\caption{A flowchart of the whole MGRFE procedure in this study.}
		\label{Fig:MGRFE_whole}
	\end{figure*}

@@

	This study involves 19 benchmark microarrays including binary, multiclass, balanced, and imbalanced datasets, which are subdivided into two large datasets.
	Dataset One consists of the 17 binary classification datasets used in ref. \cite{W16}, which includes diffuse large B-cell lymphoma (DLBCL) \cite{Y22}, Prostate (Pros) \cite{Y23}, acute lymphoblastic leukemia (ALL; subdivided into four subtypes based on different phenotypes) \cite{Y24}, central nervous system embryonal tumor (CNS) \cite{Y25}, Lymphoma (Lym) \cite{Y26}, Adenoma (Adeno) \cite{Y27}, Colon \cite{Y20}, Leukaemia (Leuk) \cite{Y21}, Myeloma (Mye)\cite{Y28}, Gastric (Gas) \cite{Y29}, and Gastric1/Gastric2 (Gas1/Gas2) cancer \cite{Y30} as well as type 1 diabetes (T1D) \cite{Y31}, and Stroke \cite{Y32}. Among them, DLBCL, Colon, Leukaemia, Myeloma, ALL1-4, and CNS datasets are imbalanced.
	Dataset Two is composed of the three typical benchmark microarray datasets used in ref. \cite{W15}, including two multiclass datasets of small-round blue-cell tumor (SRBCT) \cite{P13} and mixed lineage leukemia (MLL) \cite{W3} and one binary dataset of acute lymphoblastic leukemia and acute myeloid (ALL\_AML). Many previous experiments are conducted on these three datasets \cite{P8,P9,P10,P11,P12,P13,P14,P15,P16,P17,P18,P19,P20,P21,P22,P23}. The performance comparison between these methods and MGRFE will be provided in the Results section. It should be mentioned that the widely used benchmark Leukaemia was tested in both refs. \cite{W16} and \cite{W15} named as Leuk and ALL\_AML respectively, but they are same actually.
	The information of the 19 datasets in the two large datasets is given in Tables 1 and 2 in Supplementary Material. All these datasets can be downloaded directly from \url{https://github.com/Pengeace/MGRFE-GaRFE}.

@@

	On the 17 binary classification datasets, we employed six widely used measurements to compare the performance: Accuracy ($Acc$), Sensitivity ($Sn$), Specificity ($Sp$), Average accuracy ($Avc$), Matthews Correlation Coefficient($MCC$), and $AUC$. $AUC$ is the area under the receiver operating characteristic (ROC) curve, and the formulas of the other five measurements are presented in Equation (\ref{eq:1}). In Equation (\ref{eq:1}), $P$ and $N$ represent the numbers of positive and negative samples; $TP$ and $TN$ mean the number of correctly predicted positive and negative samples; and $FP$ and $FN$ denote the wrongly predicted positive and negative samples, respectively. For the two multiclass datasets, for consistency and convenience, only \(Acc\) is used.

@@

	As shown in Fig. \ref {Fig:MGRFE_flowchart}, the proposed MGRFE is divided into three stages, which are (1) Search space reduction; (2) Precise wrapper search and (3) Multiple \textit{k}-fold cross-validation (CV). MGRFE is a multilayer iterative feature selection method with GA-RFE acting as the feature selection unit in every layer. GA-RFE is a recursive feature elimination process including embedded GA. The Pseudocodes 1, 2 and 3 in Supplementary Material provide the pseudocodes of the processes MGRFE, GA-RFE and embedded GA.

	\begin{figure*}[t]
		\centering
		\includegraphics[width=0.87\textwidth]{pictures/MGRFE-flowchart}
		\caption{The flowchart of MGRFE, which is divided into 3 stages: search space reduction, precise wrapper search, and multiple \textit{k}-fold CV. Stage 2 is the core of MGRFE, which includes two key processes: GA-RFE and embedded modified GA.}
		\label{Fig:MGRFE_flowchart}
	\end{figure*}

@@

	At Stage 1, two filter methods, the \textit{t}-test and MIC, are used to decrease the gene range and offer a candidate gene set for later precise wrapper search stage. First, we perform the \textit{t}-test on all genes and subject them to ascending sorting according to their $p$ values, then the top-ranked statistically significant features with $p$ values less than 0.05 are preserved. Next, the upper limit of the features kept after the \textit{t}-test is set to 1000, that is, when there are more than 1000 features having $p$ values less than 0.05, only the top 1000 with lower $p$ values would be kept. If the preserved features after the \textit{t}-test screening are fewer than 500, they are all kept directly and definitively to form the candidate gene set without MIC screening; otherwise, the MIC-based selection will be followed. Second, we carry out MIC calculation on the preserved genes and resort them according to their MIC values, then the candidate gene set is generated from the top 500 genes with higher MIC values. For the two multiclass datasets, a candidate gene set is generated based only on the descending order of MIC values of all genes for the multivariate \textit{t}-test cannot be performed directly. 
	In the Table 5 in Supplementary Material, the number of statistically significant genes with \emph{t}-test-based \emph{p}-values less than 0.05 on each binary-class dataset are listed.
	In the "S8" section of Supplementary Material, we also give a simple comparison of \textit{t}-test+MIC with other filter combinations.

@@

	\subsubsection{Stage 3: Multiple \textit{k}-fold CV to select the final gene combination}

	Stage 3 is aimed at finding the optimal gene combination with the best classification performance and minimal variance among different CV processes. The \textit{k}-fold CV is used for calculating the fitness of a GA individual. Multiple \textit{k}-fold CV based on different random seeds is performed to further validate and select the final optimal gene combination.

@@

	\begin{table*}[!htbp] \footnotesize
		\centering
		\caption{Results of MGRFE on 17 datasets in Dataset One}
		\label{Tab:RST all index}
		\resizebox{0.95\textwidth}{!}{
			\begin{tabular}{l|lllllllll}
				\toprule
				Datasets & Pos/Neg & Genes/Total & $Sn$ & $Sp$ & $Acc$ & $Avc$ & $MCC$ & $AUC$ & \textit{t}-test/MIC-based gene rankings\\
				\midrule
				DLBCL & 58/19 & \textbf{3}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [13/8, 39/24, 54/52]\\
				Pros & 52/50 & \textbf{4}/12625 & 0.980 & 0.982 & \textbf{0.981} & 0.981 & 0.963 & 0.98 & [1/1, 15/47, 74/49, 694/618]\\
				Colon & 40/22 & \textbf{6}/2000 & 1.0 & 0.960 & \textbf{0.985} & 0.980 & 0.969 & 0.97 & [15/6, 58/21, 176/297, 225/80, 240/555, 495/482]\\
				Leuk & 47/25 & \textbf{2}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [4/3, 7/5]\\
				Mye & 137/36 & \textbf{7}/12625 & 0.963 & 0.839 & \textbf{0.937} & 0.901 & 0.816 & 0.95 & [3/3, 15/103, 83/142, 143/13, 378/217, 404/644, 569/707]\\
				ALL1 & 95/33 & \textbf{1}/12625 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [1/1]\\
				ALL2 & 65/35 & \textbf{8}/12625 & 0.914 & 0.908 & \textbf{0.910} & 0.911 & 0.829 & 0.94 & [1/80, 52/395, 78/3040, 80/1297, 522/2448, 687/2038, 737/920, 760/1449]\\
				ALL3 & 24/101 & \textbf{8}/12625 & 0.830 & 0.950 & \textbf{0.927} & 0.890 & 0.785 & 0.93 & [4/500, 52/3437, 75/3010, 142/393, 488/443, 510/795, 715/1551, 770/1321]\\
				ALL4 & 26/67 & \textbf{6}/12625 & 1.0 & 0.986 & \textbf{0.990} & 0.993 & 0.978 & 0.99 & [1/2, 6/45, 39/356, 282/226, 535/497, 754/1377]\\
				CNS & 39/21 & \textbf{7}/7129 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 0.98 & [9/907, 53/542, 130/620, 131/519, 272/57, 273/454, 520/49]\\
				Lym & 22/23 & \textbf{3}/4026 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [4/7, 5/4, 669/135]\\
				Adeno & 18/18 & \textbf{1}/7457 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [468/27]\\
				Gas & 29/36 & \textbf{3}/22645 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [22/1, 77/32, 306/36]\\
				Gas1 & 72/72 & \textbf{3}/22283 & 0.986 & 0.973 & \textbf{0.980} & 0.980 & 0.961 & 0.99 & [132/74, 248/167, 717/500]\\
				Gas2 & 62/62 & \textbf{2}/22283 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [38/6, 89/62]\\
				T1D & 57/44 & \textbf{7}/54675 & 0.911 & 0.912 & \textbf{0.911} & 0.912 & 0.826 & 0.94 & [14/2229, 25/1579, 113/1287, 559/1282, 578/353, 680/426, 978/1728]\\
				Stroke & 20/20 & \textbf{4}/54675 & 1.0 & 1.0 & \textbf{1.0} & 1.0 & 1.0 & 1.0 & [1/3, 23/115, 129/543, 276/539]\\
				\bottomrule
		\end{tabular}}
	\end{table*}
	\begin{figure*}[htbp]
		\centering
		\includegraphics[width=0.72\textwidth]{pictures/CV1}
		\caption{The distribution of $Acc$ values in 10-time 10-fold CV for the selected gene combinations of 17 datasets in Dataset One.}
		\label{Fig:CV1}
	\end{figure*}

@@

	McTwo \cite{W16} thoroughly tested all the datasets in Dataset One and demonstrated satisfactory performance. Here, we present the performance comparison between McTwo and MGRFE. Table \ref{Tab:CMP McTwo RST} lists the overall maximal \(Acc\) and numbers of selected genes on total 17 datasets for MGRFE and McTwo. 
	On all 17 datasets, MGRFE can obtain equal or better prediction performance compared to McTwo.	
	On five datasets ALL2, ALL3, ALL4, Stroke and CNS, MGRFE achieves distinctly better classification performance than McTwo with relatively more genes. For a fairer and more specific comparison, the \(Acc\) values of the two algorithms are listed when the gene number of MGRFE is equal to McTwo as shown in Table \ref{Tab:same gene num}. The results indicate that MGRFE still outperforms McTwo. Nonetheless, the \(Acc\) values associated with the usage of the gene numbers fall behind our optimal \(Acc\) values on these datasets. Thus, MGRFE selected somewhat more genes to achieve the optimal results.

@@

	\begin{figure}[htbp]
		\centering
		\includegraphics[width=0.45\textwidth]{pictures/CV2}
		\caption{The distribution of $Acc$ values in 10-time 10-fold CV for the selected gene combinations in Dataset Two.}
		\label{Fig:CV2}
	\end{figure}

@@

	\subsection{Independent validation of selected features}
	We performed totally 10-group validation experiments on independent datasets to verify the generalization ability of selected gene subsets by MGRFE. The results are shown in Table \ref{Tab:vali}. For each experiment, firstly, the selected gene probe features from the first dataset were transformed into the official gene symbols; secondly, the obtained gene symbols were transformed into corresponding gene probe Ids in the second dataset; thirdly, three kinds of classifier were used to perform 10 times \textit{k}-fold cross validation using the samples and selected gene probe features on the second dataset. Particularly, no feature mapping between Gas1 and Gas2 for they are generated simultaneously and have identical feature set.

	On tested datasets GSE56315 and GSE2604 with gene features from DLBCL and ALL1 respectively, NB, SVM and RF (Random Forest) classifiers all achieved 1.0 cross validation accuracy in each test. In particular, there are only 14 samples totally on GSE2604, which means the classifiers were trained on merely about 10 samples in each 5-fold cross validation. Thus, the selected unique gene \textit{CD3D} is one ideal discrimination for B-cell acute lymphoblastic leukemia (ALL) and T-cell ALL. On GSE2685, the sample number is merely 30, and only contains one gene probe Id mapped from the \textit{LIFR}, one of three selected genes on Gastric dataset. But NB and SVM still achieved acceptable cross validation accuracies over 0.8. Except three tested datasets of GSE8511, GSE2685 and GSE8514 with samples less than 50, the prediction accuracies of three classifiers are above 0.9 on all other seven datasets in cross validation. The independent validation results proved that the selected genes features by MGRFE in each dataset have strong association with the disease phenotype and can be selected as bio-marker candidates.

	\begin{table*}[htbp]
	\centering\caption{Independent validation of selected gene features by MGRFE with 10-time $k$-fold cross validation.}
	\label{Tab:vali}
	\resizebox{0.72\textwidth}{!}{
		\begin{threeparttable}
			\label{Table-vali}
			\begin{tabular}{cccccccccc}
				\toprule
				\begin{tabular}{c}
					Feature From /\\
					\#Features
				\end{tabular}&
				\begin{tabular}{c}
					Feature Tested /\\
					\#Features
				\end{tabular}&\#Samples& Classifier& \textit{Sn} &\textit{Sp}& \textit{Acc}& \textit{Avc}& \textit{MCC}& \textit{AUC}\\
				\toprule
